---
title: "Data621 - HW1"
author: "Devin Teran, Atina Karim, Tom Hill, Amit Kapoor"
date: "2/26/2021"
output:
  html_document:
    highlight: pygments
    number_sections: no
    theme: flatly
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---

```{r include=TRUE, message=FALSE, warning=FALSE}
library(dplyr)
library(DataExplorer)
library(GGally)
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Overview
The data set contains approximately 2200 records. Each record represents a professional baseball team from the years 1871 to 2006 inclusive. Each record has the performance of the team for the given year, with all of the statistics adjusted to match the performance of a 162 game season.Below is a short description of the variables

* INDEX 				    - Identification Variable
* TARGET_WINS       - Number of wins
* TEAM_BATTING_H		- Base Hits by batters (1B,2B,3B,HR)
* TEAM_BATTING_2B		- Doubles by batters (2B)
* TEAM_BATTING_3B		- Triples by batters (3B)
* TEAM_BATTING_HR		- Homeruns by batters (4B)
* TEAM_BATTING_BB		- Walks by batters 
* TEAM_BATTING_HBP	- Batters hit by pitch (get a free base) 
* TEAM_BATTING_SO		- Strikeouts by batters 
* TEAM_BASERUN_SB		- Stolen bases 
* TEAM_BASERUN_CS		- Caught stealing 
* TEAM_FIELDING_E		- Errors 
* TEAM_FIELDING_DP	- Double Plays 
* TEAM_PITCHING_BB	- Walks allowed 
* TEAM_PITCHING_H		- Hits allowed 
* TEAM_PITCHING_HR	- Homeruns allowed 
* TEAM_PITCHING_SO	- Strikeouts by pitchers 

## Objective
To build a multiple linear regression model on the training data to predict the number of wins for the team.



## Data Exploration

```{r read-csv}
# read data
moneyball <- read.csv("https://raw.githubusercontent.com/hillt5/DATA_621/master/HW1/moneyball-training-data.csv", header=TRUE)
head(moneyball)
```

```{r}
# dimension
dim(moneyball)
```

There are 2276 rows and 17 columns in the dataset.

```{r}
# glimpse of data
glimpse(moneyball)
```


```{r}
# removing index
moneyball <- moneyball %>% dplyr::select(-INDEX)
```


```{r}
# data summary
summary(moneyball)
```



```{r}
# check NAs in columns
colSums(is.na(moneyball))
```




```{r}
# distribution
plot_histogram(moneyball)
```

```{r}
# against the response variable
plot_scatterplot(moneyball, by = "TARGET_WINS")
```


```{r}
# boxplot for train dataset
plot_boxplot(moneyball, by="TARGET_WINS")
```


```{r}
# correlation
moneyball %>% 
  ggcorr(size =3, hjust=1.0)
```





## Data Preparation

## Build Models

## Select Models








