---
title: "Data 621"
subtitle: 'Group 4 - Homework1'
author: "Devin Teran, Atina Karim, Tom Hill, Amit Kapoor"
date: "3/1/2020"
output:
  html_document: default
  pdf_document:
    toc: TRUE
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, error=FALSE, warning=FALSE, message=FALSE)
```

```{r loadData, include=FALSE}
# Libraries
library(dplyr)
library(DataExplorer)
library(GGally)
library(ggplot2)
library(readr)
library(reshape2)
library(purrr)
library(tidyr)
library(corrplot)
library(MASS)
library(caret)

set.seed(2012)

# read data
baseball_df <- read.csv('https://raw.githubusercontent.com/hillt5/DATA_621/master/HW1/moneyball-training-data.csv')
baseball_eval <- read.csv('https://raw.githubusercontent.com/hillt5/DATA_621/master/HW1/moneyball-evaluation-data.csv')


```

# DATA EXPLORATION
The dataset given contains 2276 records for baseball teams from the years 1871 to 2006. There are 15 features (predictors) for a team's performance in a season and a target (response) variable for team's wins in that season. A baseball season has 162 games. The wins distribution seems nearly normal distributed. It indicates that seasons do not have many too many high or low number of wins.


```{r histWins}
plot_histogram(baseball_df$TARGET_WINS, 
               title="Distribution of TARGET_WINS")
```




# DATA PREPARATION
Now we have good understanding of data so we can move to prepare the data for buliding model. 

## Outliers
Let's see once the outliers for each predictor.


```{r dataOutliers}
baseball_df %>% 
  dplyr::select(-INDEX, -TARGET_WINS) %>% 
  pivot_longer(everything(), names_to = 'Var', values_to='Value') %>% 
  ggplot(aes(x = Var, y = Value)) +
  geom_boxplot() + 
  coord_flip()
  
```




# BUILD MODELS



# SELECT MODELS



# REFERENCES
* An Introduction to Statistical Learning
with Applications in R
Springer


# CODE APPENDIX
The code chunks below shows the R code called above throughout the analysis.
They are being reproduced in the appendix for review and feedback.

```{r appendix, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, eval=FALSE)
```


```{r setup}
```

```{r loadData}
```

```{r histWins}
```

```{r dataOutliers}
```








