---
title: "Data 621 - HW3"
author: "Devin Teran, Atina Karim, Tom Hill, Amit Kapoor"
date: "4/18/2021"
output:
  pdf_document:
    toc: TRUE
    toc_depth: 2 
  html_document:
    highlight: pygments
    number_sections: no
    theme: flatly
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, error=FALSE, warning=FALSE, message=FALSE)
```

```{r loadData, include=FALSE}


# Libraries
library(dplyr)
library(GGally)
library(ggplot2)
library(readr)
library(reshape2)
library(purrr)
library(tidyr)
library(corrplot)
library(MASS)
library(caret)
library(e1071)

        
set.seed(2012)

crime_training <- read.csv('https://raw.githubusercontent.com/hillt5/DATA_621/master/HW3/crime-training-data_modified.csv')
crime_eval <- read.csv('https://raw.githubusercontent.com/hillt5/DATA_621/master/HW3/crime-evaluation-data_modified.csv')


```


```{r, crime-df}

head(crime_training)


```

Explanation of features:


 znn: proportion of residential land zoned for large lots (over 25000 square feet) (predictor variable)
 indus: proportion of non-retail business acres per suburb (predictor variable)
 chas: a dummy var. for whether the suburb borders the Charles River (1) or not (0) (predictor variable)
 nox: nitrogen oxides concentration (parts per 10 million) (predictor variable)
 rm: average number of rooms per dwelling (predictor variable)
 age: proportion of owner-occupied units built prior to 1940 (predictor variable)
 dis: weighted mean of distances to five Boston employment centers (predictor variable)
 rad: index of accessibility to radial highways (predictor variable)
 tax: full-value property-tax rate per $10,000 (predictor variable)
 ptratio: pupil-teacher ratio by town (predictor variable)
 black: 1000(Bk - 0.63)2 where Bk is the proportion of blacks by town (predictor variable)
 lstat: lower status of the population (percent) (predictor variable)
 medv: median value of owner-occupied homes in $1000s (predictor variable)
 target: whether the crime rate is above the median crime rate (1) or not (0) (response variable)
